---
title: "p8105_hw1_ob2305"
author: "Olya Besedina"
date: 2019-09-18
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
```

# Problem 1

```{r problem 1}

my_df = tibble(
  vec_numeric = rnorm(8),
  vec_logical = vec_numeric > 0,
  vec_char = c("monkey", "eel", "finch", "dog", "cat", "salmon", "carp","parrot"),
  vec_factor = factor(c("mammal", "fish", "bird", "mammal", "mammal", "fish", "fish", "bird"))
)

# The mean of the variables inside the data frame
mean(pull(my_df, vec_numeric))
mean(pull(my_df, vec_logical))
mean(pull(my_df, vec_char))
mean(pull(my_df, vec_factor))

# You can only take a mean of numeric or logical vectors, thus mean(vec_char) and mean(vec_factor) are not applicable.

## Application of as.numeric

# logical vector
l = pull(my_df, vec_logical)
as.numeric(l)

# character vector
c = pull(my_df, vec_char)
as.numeric(c)
# You cannot apply as.numeric function to a character vector. This explains why you cannot take a mean of words. 

# factor vector
f = pull(my_df, vec_factor)
as.numeric(f)
# Factor vector has 3 levels and thus can be expressed as a number. 
```

## Second code chunk

Convert logical vector to numeric and multiply result by the random sample
```{r eval=FALSE}
l = pull(my_df, vec_logical)
l_1 = as.numeric(l)
pull(my_df,vec_numeric)*l_1
```

Convert the logical vector to a factor and multiply teh random sample by the result
```{r eval=FALSE}
l_2 = pull(my_df, vec_logical)
l_3 = as.factor(l_2)
pull(my_df,vec_numeric)*l_3
```

Convert logical vector to a factor and then convert the result to numeric and multiply random sample by the result
```{r eval=FALSE}
l_2 = pull(my_df, vec_logical)
l_3 = as.factor(l_2)
l_4 = as.numeric(l_3)
pull(my_df,vec_numeric)*l_4
```


# Problem 2

```{r problem 2}
plot_df = tibble(
  x = rnorm(500),
  y = rnorm(500),
  logic_vector = (x + y) >1,
  numeric_vector = as.numeric(logic_vector),
  factor_vector = as.factor(logic_vector)
)

df_rows = nrow(plot_df)
df_col = ncol(plot_df)
x_mean = mean(pull(plot_df, x))
x_median = median(pull(plot_df, x))
x_sd = sd(pull(plot_df, x))
lv = pull(plot_df, logic_vector)
```
My data frame has `r df_rows` rows, `r df_col` columns. The mean of x is equal to `r round(x_mean, digits = 3)`, median = `r round(x_median, digits=3)` and standard deviation = `r round(x_sd, digits=3)`. The proportion of cases for which (x+y) > 1 is equal to `r sum(lv)`.

Here are the scatter plots
```{r yx_scatter}
#  scatter plot of x and y
ggplot(plot_df, aes(x = x, y = y,  color = logic_vector)) + geom_point()
ggsave("scatterplot.pdf")
```

Scatter plot has two colors representing true and false values for logical vector x+y > 1; values which are >1 represented by the teal color, while values which are <1 represented by the red color. 

```{r numeric_scatter}
# scatter plot for numeric variable 
ggplot(plot_df, aes(x = x, y = y,  color = numeric_vector)) + geom_point()
```

Scatter plot has two colors representing values for a numeric factor; values which are >1 represented by the light blue color, while values which are <1 represented by the navy color. 

```{r factor_scatter}
# scatter plot for factor variable 
ggplot(plot_df, aes(x = x, y = y,  color = factor_vector)) + geom_point()
```

Scatter plot has two colors representing true and false values for factor variable; values which are >1 represented by the teal color, while values which are <1 represented by the red color. 
